---
title: "Calculate and assess Zinc DGVs -Australia and New Zealand - MERA project"
author: "Jenni Gadd"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
format:
  html:
    code-fold: false    
    code-overflow: wrap    
    toc: true
    toc-expand: 3
    toc-location: left
    toc-title: Table of contents
    number-sections: true
editor: visual
execute: 
  cache: false
  warning: false
  echo: false
editor_options: 
  chunk_output_type: console
---

Packages used for the analysis.

```{r setup}
#|output: false 
#source("ZnZnDGVscript_ssdtools.R")  

library(ggplot2)
library(ggpubr)
library(ggridges)
library(cowplot)
library(tidyr)
library(dplyr)
library(lubridate)
library(readxl)
library(writexl)
library(tidyxl)  # reading in untidy Excel spreadsheets
library(readr)
library(skimr)
library(flextable)
set_flextable_defaults(big.mark = "")

library(tidyverse)

# nzmap <- rnaturalearth::ne_countries(
#   scale = "large",
#   country = "New Zealand",
#   returnclass = "sf"
# )
# ## for stadia maps type------
# register_stadiamaps("f6a0c48d-2743-48b6-bd5c-85be122b1251", write = FALSE)
# 
# bbox <- c(164.81, -47.74, 179.44, -33.79)
# names(bbox) <- c("left", "bottom", "right", "top")
# 
# nz.terrain <- ggmap::get_stadiamap(bbox, maptype = "stamen_terrain_background", zoom = 7)

#nzborders <- st_transform(nz, crs = 4326)
```

# Background

All the TMF data and metal concentrations (if measured) are found in a wide-style dataframe, with columns that indicate State, site name, site code and censoring.

The data frame includes all years and we can crop it later to more recent data.

# Import the dataframe and quick checks

```{r}
## change this link to the tmf data a
# load(file = "P:\\231220_ANZG_MERA_Imp\\Task 1_Data collation\\6.Australia\\output\\aus.data.wide_updated.Rdata")
#  mydata <- aus.data.wide
# # 
#  names(mydata)
#  skim(mydata)
# # 
# mydata$DOC <- mydata$`Dissolved organic carbon`
#  mydata$Ca <- mydata$`Calcium (soluble)`
#  mydata$Mg <- mydata$`Magnesium (soluble)`
#  mydata$Calcium <- mydata$`Calcium (soluble)`
#  mydata$Magnesium <- mydata$`Magnesium (soluble)`
#  mydata$H <- mydata$`Hardness`  # this one is calculated where H was missing
#  mydata <- mydata |> dplyr::mutate(row = row_number(),
#  .before = State)
# # 
#  mydata.snip <- head(mydata, 10)
#  mydata$Country <- "Australia"
# # 
#  mydata.snip
#  
#  # # New Zealand
#  load(file = "output/tmf.data.wide_updated.Rdata")
# nz_doc <- tmf.data.wide |>
#   filter(DOC > 0)
# 
# save(nz_doc,  file = "output/nz.doc.for.ni.Rdata")
```

## Some info- Aus

```{r}
# 
# length(!is.na(aus.data.wide$DOC))
# sum(aus.data.wide$DOC > 0, na.rm =TRUE)
# sum(aus.data.wide$DOC > 0 & aus.data.wide$DOC <=30, na.rm =TRUE)
# 
# # ## Check censoring of inputs
# badDOC <- aus.data.wide |>
#   dplyr::filter(`Prefix_Dissolved.Organic.Carbon` == "<"& DOC >= 1)
# 
# nrow(badDOC)
# 
# aus_doc <- mydata |>
#   select(-`Prefix_zinc (dissolved)`, -`Prefix_zinc (total)`, -`Prefix_Zinc (dissolved)`, -`Prefix_Zinc (total)`, -`Prefix_zinc (total)`, -`Prefix_Total Magnesium`, -`Prefix_Total Organic Carbon (TOC)`, 
#          -`Prefix_Total carbon`, -`Prefix_Field pH`, -Prefix_pH, -Prefix_zinc.dissolved.mg.L, -Prefix_zinc.dissolved.mg.L, -Prefix_Zinc.dissolved..mg.L., -`Prefix_pH (field)`, -n_param, -DOCTOC, -SURVEYCODE, -DATETIME, -id, -row,   -lH,
#          -lCa, -lMg, -`Dissolved zinc`, -`Dissolved zinc`, 
#          -zinc.Total.mg.L, -zinc.dissolved.mg.L, -zinc.Total..mg.L, -Zinc.Total.mg.L, -Total.Organic.Carbon, -`Total organic carbon`, -`Total carbon`,
#          -Check_H, -Source, -Description, -DEPTH, -TIME, -BOTTLE, -filename, -CaMgratio, -Sampling.Depth..m., -`Prefix_Total organic carbon`) |>
#   
#   filter(DOC>0)
# write.csv(aus_doc, file = "output/aus.doc.for.ni.csv")
# save(aus_doc, file = "output/aus.doc.for.ni.Rdata")

# aussie_tmfs <- aussie_gvs |> select(-c(Ni_BAGV99: NiNote))
# 
# write.csv(aussie_tmfs, file = "P:/240209_Zinc_freshwater/8. BAGV_DGV calcs/TMFdf_Aus.csv")

```

```{r somefixesfirst}

# load(file = "P:\\231220_ANZG_MERA_Imp\\Task 1_Data collation\\3.DataProcessing\\output\\tables\\TMFdfwithGVs.Rdata")
# 
# AllMetalGVs <- AllMetalGVs |> mutate(DOC = ifelse(DOC <0.2, 0.2, DOC),
#                                      Hardness = H) |>
#   select(-c(CuPC99: NiPC80)) 
# 
# NZ_df <- AllMetalGVs
# save(NZ_df, file = "P:/240209_Zinc_freshwater/8. BAGV_DGV calcs/TMFdf_NZ.Rdata")
# write.csv(NZ_df, file = "P:/240209_Zinc_freshwater/8. BAGV_DGV calcs/TMFdf_NZ.csv")
```

## IMPORT AUS & NZ values with Zn BAPCs

```{r importnz}
##NZ
#load(file = "output/nz.doc.for.ni.Rdata")
# Change path to whereever the file is--------------
# load(file = "P:\\231220_ANZG_MERA_Imp\\Task 1_Data collation\\3.DataProcessing\\output\\tables\\TMFdfwithGVs.Rdata")
load(file = "data\\BAGVs for Australia and New Zealand\\TMFdfwithGVs.Rdata")
#
# write.csv(AllMetalGVs, file = "P:/240209_Zinc_freshwater/8. BAGV_DGV calcs/TMFdf_NZ.csv")

nz_zn <- AllMetalGVs |># dplyr::filter(ZnPC99 >0)
 dplyr::filter(Date >= "2000-01-01") |>
     dplyr::mutate(Hardness = `Hardness..total.`) |>
  rename(ZnPC99 = Zn_BAGV99, ZnPC95 = Zn_BAGV95, 
           ZnPC90  = Zn_BAGV90, ZnPC80 = Zn_BAGV80) |>
  relocate(ZnPC99, ZnPC95, ZnPC90, ZnPC80, Hardness, .after = DOC)


## Remove thos with high censoring
badDOC <- nz_zn |>
  dplyr::filter(`Censored_Dissolved.organic.carbon` == "TRUE" &  DOC >= 1)

nrow(badDOC)

nz_zn <- nz_zn |>
  #dplyr::filter(!(`Prefix_Dissolved.Organic.Carbon` == "<" & DOC >= 1))
dplyr::filter(!(row %in% badDOC$row))
```

```{r importaus}

## Change this link
#load(file = "output/aus.doc.for.ni.Rdata")
# aussie_gvs <- read_xlsx(path = "P:\\231220_ANZG_MERA_Imp\\Task 2_CreateRefGVs\\MyDGVs-2025-03-10_Aus.xlsx", sheet = "gvs")
aussie_gvs <- read_excel(path = "data\\BAGVs for Australia and New Zealand\\Aus-DGVs-2025-08-12.xlsx", sheet = "gvs", guess_max=2000)

aussie_zn <- aussie_gvs  |> 
  dplyr::filter(Date >= "2000-01-01") |>
     dplyr::select(-Zn_DGV99,-Zn_DGV95,-Zn_DGV90, -Zn_DGV80) |>
    rename(ZnPC99 = Zn_BAGV99, ZnPC95 = Zn_BAGV95, 
           ZnPC90  = Zn_BAGV90, ZnPC80 = Zn_BAGV80) |>
#   #dplyr::select(-CuPC99, -CuPC95, -CuPC90, -CuPC80) |> 
    relocate(ZnPC99, ZnPC95, ZnPC90, ZnPC80, Hardness, DOC, .after = pH) |> 
  dplyr::filter(ZnPC99 >0)

## Remove thos with high censoring
badDOC <- aussie_zn |>
  dplyr::filter(`Prefix_Dissolved.Organic.Carbon` == "<" &  DOC >= 1)

nrow(badDOC)

aussie_zn <- aussie_zn |> 
  #dplyr::filter(!(`Prefix_Dissolved.Organic.Carbon` == "<" & DOC >= 1))
dplyr::filter(!(row %in% badDOC$row))
```

### Summaries for zinc - all BAGVs

```{r znsummary-aus}

quantile(aussie_zn$ZnPC95, probs= c(0.01, 0.99), na.rm = TRUE)

ZnBAGV.sum <- aussie_zn |> 
 dplyr::summarise(across(
    .cols = contains("PC"),
    .fns =  list(`No.PCs calcd` = ~sum(!is.na(.)),
                 Quantile1 = ~quantile(., probs = 0.01, na.rm = TRUE),
                 Quantile2 = ~quantile(., probs = 0.02, na.rm = TRUE),
                 Quantile5 = ~quantile(., probs = 0.05, na.rm = TRUE),
                 Quantile10 = ~quantile(., probs = 0.1, na.rm = TRUE),
                 MedianPC = ~median(., na.rm = TRUE),
                 MinPC = ~min(., na.rm = TRUE),
                 MaxPC = ~max(., na.rm = TRUE)
                 ))) %>%
    pivot_longer(cols = everything(), 
               names_pattern = "(.*)_(.*)", 
             names_to = c("group", ".value")) 

ZnBAGV.sum2 <- ZnBAGV.sum  %>% 
  dplyr::mutate(across(Quantile1:MaxPC, \(x) signif (x, digits=2))) |>
  column_to_rownames(var="group") |>
  t() |>
  as.data.frame() |>
  rownames_to_column(var="Statistic")
 
grey.index <- as.logical(c(FALSE, rep(TRUE,4)))

ZnBAGV.sum2 |>
  flextable() |>
  add_header_lines("Summary of Zn BAGVs") |>
  autofit() |>
   bg(j = grey.index, bg = "grey90") |>
   bg(i= ~ Statistic == "Quantile5", bg = "pink")

summary(aussie_zn$ZnPC95)
```

```{r zn-summary-nz}
quantile(nz_zn$ZnPC95, probs= c(0.01, 0.99), na.rm = TRUE)

ZnBAGV.sum <- nz_zn |> 
  #dplyr::select(-DGV_ZnPC99, -DGV_ZnPC95, -DGV_ZnPC90, -DGV_ZnPC80) |>
  dplyr::summarise(across(
    .cols = contains("PC"),
    .fns =  list(`No.PCs calcd` = ~sum(!is.na(.)),
                 Quantile1 = ~quantile(., probs = 0.01, na.rm = TRUE),
                 Quantile2 = ~quantile(., probs = 0.02, na.rm = TRUE),
                 Quantile5 = ~quantile(., probs = 0.05, na.rm = TRUE),
                 Quantile10 = ~quantile(., probs = 0.1, na.rm = TRUE),
                 MedianPC = ~median(., na.rm = TRUE),
                 MinPC = ~min(., na.rm = TRUE),
                 MaxPC = ~max(., na.rm = TRUE)
                 ))) %>%
    pivot_longer(cols = everything(), 
               names_pattern = "(.*)_(.*)", 
             names_to = c("group", ".value")) 

ZnBAGV.sum2 <- ZnBAGV.sum  %>% 
  dplyr::mutate(across(Quantile1:MaxPC, \(x) signif (x, digits=2))) |>
  column_to_rownames(var="group") |>
  t() |>
  as.data.frame() |>
  rownames_to_column(var="Statistic")
 
grey.index <- as.logical(c(FALSE, rep(TRUE,4)))

ZnBAGV.sum2 |>
  flextable() |>
  add_header_lines("Summary of Zn BAGVs") |>
  autofit() |>
  # bg(j = grey.index, bg = "grey90") |>
   bg(i= ~ Statistic == "Quantile5", bg = "pink")

# write.csv(ZnBAGV.sum2, file = "output/nz_zn_stats.csv")
write.csv(ZnBAGV.sum2, file = "data/BAGVs for Australia and New Zealand/output/nz_zn_stats.csv")


BAGV.sum2  <- ZnBAGV.sum2  %>% 
  dplyr::select(Statistic, contains('PC95'))
  
BAGV.sum2 |>
  flextable() |>
  add_header_lines("Summary of Zn BAGVs") |>
  autofit() |>
  bg(i= ~ Statistic == "Quantile5", bg = "pink")    

```

## Filter to values in range

```{r}
# Split into 2 DFs - one for zinc and one for zinc

ausZnGVs_definrange <- aussie_zn |> 
  mutate(Country = "Australia") |>
  filter(pH >= 5.99999 & pH < 8.1) |>
  filter(Calcium >= 3.7 & Calcium <= 88.4)|>
  filter(Magnesium >= 3 & Magnesium <= 72.4)|>
  filter(DOC >= 0.5 & DOC <= 17.4)

summary(ausZnGVs_definrange$ZnPC95)
nrow(ausZnGVs_definrange)
quantile(ausZnGVs_definrange$ZnPC95, probs= c(0.01, 0.99), na.rm = TRUE)
# 
# pHout <- aussie_zn |> 
#   filter(!is.na(ZnPC95)) |>
#   filter(!is.na(pH) & !is.na(H)) |>
#   filter(pH < 5.99999 |  pH >= 8.01 )
# 
# Hout <- aussie_zn |> 
#   filter(!is.na(ZnPC95)) |>
#   filter(!is.na(pH) & !is.na(H)) |>
#   filter(H<2 |  H > 340 )
# 
# DOCout <- aussie_zn |> 
#   filter(!is.na(ZnPC95)) |>
#   filter(!is.na(pH) & !is.na(H)) |>
#   filter(DOC >17) 
  
nzZnGVs_definrange <- nz_zn |> 
  dplyr::mutate(Country = "New Zealand") |>
  filter(pH >= 5.99999 & pH < 8.1) |>
  filter(Ca >= 3.7 & Ca <= 88.4)|>
  filter(Mg >= 3 & Mg <= 72.4)|>
  filter(DOC >= 0.5 & DOC <= 17.4)

nrow(nzZnGVs_definrange)

summary(nzZnGVs_definrange$ZnPC95)
quantile(nzZnGVs_definrange$ZnPC95, probs= c(0.01, 0.99), na.rm = TRUE)
```

## Compare in range & out of range

```{r compare_inout}
##  zinc 
zn.combo.GV.data <- rbind(ausZnGVs_definrange  |> 
                         mutate(Dataset = "In range"),
                       aussie_zn |> filter(!is.na(ZnPC95)) |> 
                         mutate(Dataset = "All calculated GVs")) 

Zn_summary_stats <- zn.combo.GV.data |> 
  dplyr::group_by(Dataset) |>
  dplyr::summarise(Median = median(ZnPC95),
                   Min =min(ZnPC95),
                   Max = max(ZnPC95),
                   Q1 = quantile(ZnPC95, probs = 0.01),
                   Q5 = quantile(ZnPC95, probs = 0.05)) |>
  dplyr::mutate(Y = c(1.5, 2.5))

Zn_summary_stats |>
  dplyr::select(-Y, -Q1) |>
  flextable() |>
  add_header_lines("Summary of zinc BAGVs in & out of range") |>
  autofit() |>
  theme_zebra()

ggplot(data = zn.combo.GV.data,
       aes(x = ZnPC95, y = Dataset, fill = Dataset)) +
    stat_density_ridges(quantile_lines = TRUE, quantiles = c(0.05), 
                         linetype = 3, alpha =0.5) +
  geom_density_ridges(alpha = 0, 
                      jittered_points = TRUE, position = position_points_jitter(width = 0.05, height = 0),
                       point_shape = '|', point_size = 1, point_alpha = 1  ) +
    scale_y_discrete(expand = c(0, 0)) +     # will generally have to set the `expand` option
  scale_x_log10(expand = c(0, 0)#, limits = c(0,40)
                     ) +   # for both axes to remove unneeded padding
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  
  annotate("text", x = 0, y = Zn_summary_stats$Y, hjust = 1.1, vjust = 1.1,
             label = paste("5th %ile:", round(Zn_summary_stats$Q5, 1)#, "\n1st %ile:", round(Zn_summary_stats$Q1, 2)
                           )) +
  theme_ridges() +
  ylab(NULL)+
  xlab(expression("zinc BAPC95 (95% species protection, "*mu*"g/L)")) +
  theme(legend.position = "none")


# ggsave(file = "output/figuresnew/ZnGVs_invsout_aus_feb25.png", width = 15, height = 12, units = "cm")
ggsave(file = "data/BAGVs for Australia and New Zealand/output/ZnGVs_invsout_aus_aug25.png", width = 15, height = 12, units = "cm")

## NZ-------------

zn.combo.GV.data <- rbind(nzZnGVs_definrange  |> 
                         mutate(Dataset = "In range"),
                       nz_zn |> filter(!is.na(ZnPC95)) |> 
                         mutate(Dataset = "All calculated GVs")) 

Zn_summary_stats <- zn.combo.GV.data |> 
  dplyr::group_by(Dataset) |>
  dplyr::summarise(Q1 = quantile(ZnPC95, probs = 0.01, na.rm=TRUE),
                   Q5 = quantile(ZnPC95, probs = 0.05, na.rm=TRUE)) |>
  dplyr::mutate(Y = c(1.5, 2.5))

Zn_summary_stats |>
 # select(-Y, -Q1) |>
  flextable() |>
  add_header_lines("Summary of zinc BAGVs in & out of range") |>
  autofit() |>
  theme_zebra()

ggplot(data = zn.combo.GV.data,
       aes(x = ZnPC95, y = Dataset, fill = Dataset)) +
    stat_density_ridges(quantile_lines = TRUE, quantiles = c(0.05), 
                         linetype = 3, alpha =0.5) +
  geom_density_ridges(alpha = 0, 
                      jittered_points = TRUE, position = position_points_jitter(width = 0.05, height = 0),
                       point_shape = '|', point_size = 1, point_alpha = 1  ) +
    scale_y_discrete(expand = c(0, 0)) +     # will generally have to set the `expand` option
  scale_x_log10(expand = c(0, 0)#, limits = c(0,40)
                     ) +   # for both axes to remove unneeded padding
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  
  annotate("text", x = 0, y = Zn_summary_stats$Y, hjust = 1.1, vjust = 1.1,
             label = paste("5th %ile:", round(Zn_summary_stats$Q5, 1)#, "\n1st %ile:", round(Zn_summary_stats$Q1, 2)
                           )) +
  theme_ridges() +
  ylab(NULL)+
  xlab(expression("zinc BAPC95 (95% species protection, "*mu*"g/L)")) +
  theme(legend.position = "none")


# ggsave(file = "figuresnew/ZnGVs_invsout_nz_feb25.png", width = 15, height = 12, units = "cm")
ggsave(file = "data/BAGVs for Australia and New Zealand/output/ZnGVs_invsout_nz_aug25.png", width = 15, height = 12, units = "cm")
```

# Compare countries

```{r coutnryq5tables}

nzZnQ1_5s  <- nzZnGVs_definrange |>
     dplyr::summarise(N.pts = sum(!is.na(ZnPC95)),
                       across(
    .cols = c(ZnPC99: ZnPC80),
    .fns = list(
 Min =  ~min(.,na.rm = TRUE),
   Quantile5 = ~quantile(., probs = 0.05, na.rm = TRUE),
     Quantile25 = ~quantile(., probs = 0.25, na.rm = TRUE),
   Median = ~quantile(., probs = 0.5, na.rm = TRUE),
     Quantile75 = ~quantile(., probs = 0.75, na.rm = TRUE),
   Quantile95 = ~quantile(., probs = 0.95, na.rm = TRUE),
   Max = ~max(.,na.rm = TRUE)
))) |>
  pivot_longer(
    cols = everything(),
    names_to = c("Statistic", "PC"),
    names_sep = "_",
    values_to = "Value") |>
    pivot_wider(names_from = "Statistic", values_from= "Value")
  

nzZnQ1_5s |>
  flextable() |>
  add_header_lines("Summary of BAGVs - NZ ") |>
  autofit() |>
  theme_zebra()

## Aussie

ausZnQ1_5s  <- ausZnGVs_definrange |>
     dplyr::summarise(N.pts = sum(!is.na(ZnPC95)),
                       across(
    .cols = c(ZnPC99: ZnPC80),
    .fns = list(
  Min =  ~min(.,na.rm = TRUE),
   Quantile5 = ~quantile(., probs = 0.05, na.rm = TRUE),
     Quantile25 = ~quantile(., probs = 0.25, na.rm = TRUE),
   Median = ~quantile(., probs = 0.5, na.rm = TRUE),
     Quantile75 = ~quantile(., probs = 0.75, na.rm = TRUE),
   Quantile95 = ~quantile(., probs = 0.95, na.rm = TRUE),
   Max = ~max(.,na.rm = TRUE)
))) |>
  pivot_longer(
    cols = everything(),
    names_to = c("Statistic", "PC"),
    names_sep = "_",
    values_to = "Value") |>
    pivot_wider(names_from = "Statistic", values_from= "Value")
  
ausZnQ1_5s |>
  flextable() |>
  add_header_lines("Summary of BAGVs Aus ") |>
  autofit() |>
  theme_zebra()

```

```{r}

ggplot(nzZnGVs_definrange, aes(ZnPC95)) +
  stat_ecdf(geom = "step") +
  theme_bw() +
  xlab(expression("zinc BAGV ("*mu*"g/L)")) +
  ylab("Percentage GVs")

# rbind(nzZnGVs_definrange |> select(Country, ZnPC95), ausZnGVs_definrange |> select(Country, ZnPC95)) |>
# ggplot() +
#   geom_boxplot(aes(x= Country, y =ZnPC95), width =0.25) +
#   theme_bw() +
#   ylab(expression("zinc BAPC95 ("*mu*"g/L)")) +
#   xlab(NULL)

# rbind(nzZnGVs_definrange |> dplyr::select(Country, ZnPC95), 
#       ausZnGVs_definrange |> 
#         dplyr::select(Country, ZnPC95)) |>
# ggplot() +
#   geom_jitter(aes(x= Country, y =ZnPC95), width = 0.1, alpha = 0.2, colour = "#011D77") +
#   geom_boxplot(aes(x= Country, y =ZnPC95), width = 0.2, outliers = FALSE, 
#                colour = "#011D77", fill = "transparent") +
#     theme_bw() +
#   ylab(expression("zinc BAPC95 (95% species protection, "*mu*"g/L)")) +
#   xlab(NULL)

Zn_summary_stats <- rbind(nzZnGVs_definrange |> dplyr::select(Country, ZnPC95), 
                          ausZnGVs_definrange |> dplyr::select(Country, ZnPC95)) |> 
  dplyr::group_by(Country) |>
  dplyr::summarise(Q5 = quantile(ZnPC95, probs = 0.05, na.rm=TRUE)) |>
  dplyr::mutate(Y = c(1.5, 2.5))

znp <-rbind(nzZnGVs_definrange |> dplyr::select(Country, ZnPC95), 
      ausZnGVs_definrange |> dplyr::select(Country, ZnPC95)) |>

ggplot(
       aes(x = ZnPC95, y = Country, fill = Country)) +
    stat_density_ridges(quantile_lines = TRUE, quantiles = c(0.05), 
                         linetype = 3, alpha =0.5) +
  geom_density_ridges(alpha = 0, 
                      jittered_points = TRUE, position = position_points_jitter(width = 0.05, height = 0),
                       point_shape = '|', point_size = 1, point_alpha = 1  ) +
    scale_y_discrete(expand = c(0, 0)) +     # will generally have to set the `expand` option
  scale_x_log10(expand = c(0, 0)#, limits = c(0,40)
                     ) +   # for both axes to remove unneeded padding
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  
  annotate("text", x = 1, y = 1.05*Zn_summary_stats$Y, hjust = 1, vjust = 1.1,
             label = paste("5th %ile:", round(Zn_summary_stats$Q5, 1)#, "\n1st %ile:", round(Zn_summary_stats$Q1, 2)
                           )) +
  theme_ridges() +
  ylab(NULL)+
  xlab(expression("zinc BAPC95 ("*mu*"g/L)")) +
  theme(legend.position = "none") + 
  scale_fill_manual(values = c("New Zealand" = "grey40", "Australia" = "goldenrod"))


znp

#ggsave(file = "output/figuresnew/ZnGVs_ausvnz.png", width = 20, height = 12, units = "cm")
ggsave(file = "data/BAGVs for Australia and New Zealand/output/ZnGVs_ausvnz.png", width = 20, height = 12, units = "cm")


Zn_summary_stats <- rbind(nzZnGVs_definrange |> dplyr::select(Country, DOC), 
                          ausZnGVs_definrange |> dplyr::select(Country, DOC)) |> 
  dplyr::group_by(Country) |>
  dplyr::summarise(Q5 = quantile(DOC, probs = 0.05)) |>
  dplyr::mutate(Y = c(1.5, 2.5))



```

# zinc

Plot the outputs for zinc - focusing on BAGVs in range and PC 95

### Overall plots

```{r znsums}

znGV.summary <- ausZnGVs_definrange |>
  
  ungroup() |>
  filter(!is.na(ZnPC95)) |>
  group_by(State) |>
   #group_by(Region) |>
  dplyr::summarise(
   Min = min(ZnPC95, na.rm = TRUE),
    Q5 = quantile(ZnPC95, probs = 0.05, na.rm = TRUE),
    Q25 = quantile(ZnPC95, probs = 0.25, na.rm = TRUE),
     Median = median(ZnPC95, na.rm = TRUE),
    Q75 = quantile(ZnPC95, probs = 0.75, na.rm = TRUE),
    Q95 = quantile(ZnPC95, probs = 0.95, na.rm = TRUE),
   
    Max = max(ZnPC95, na.rm = TRUE),
    N = n()
  ) |>
  arrange(Min)

znGV.summary |>
  flextable() |>
  autofit() |>
  theme_zebra()


znGV.summary <- ausZnGVs_definrange |>
   ungroup() |>
  filter(!is.na(ZnPC95)) |>
    dplyr::summarise(
   Min = min(ZnPC95, na.rm = TRUE),
    Q5 = quantile(ZnPC95, probs = 0.05, na.rm = TRUE),
    Q25 = quantile(ZnPC95, probs = 0.25, na.rm = TRUE),
     Median = median(ZnPC95, na.rm = TRUE),
    Q75 = quantile(ZnPC95, probs = 0.75, na.rm = TRUE),
    Q95 = quantile(ZnPC95, probs = 0.95, na.rm = TRUE),
   
    Max = max(ZnPC95, na.rm = TRUE),
    N = n()
  ) |>
  arrange(Median)


znGV.summary |>
  flextable() |>
  autofit() |>
  add_header_lines("Summary of BAGVs Aus ") |>
 theme_zebra()

ggplot(ausZnGVs_definrange) +
  geom_boxplot(aes(x = reorder(State, ZnPC95, FUN = median), y= ZnPC95)) + 
    # geom_jitter(aes(x = reorder(State, ZnPC95, FUN = median), y= ZnPC95),
    #             width = 0.2, alpha = 0.1) + 
   xlab(NULL) +
   ylab(expression("zinc BAPC95 (95% species protection, "*mu*"g/L)")) +
   theme_bw()

# ggsave(file = "output/figuresnew/ZnGVs_boxaus_feb.png", width = 20, height = 12, units = "cm")
 
ggsave(file = "data/BAGVs for Australia and New Zealand/output/ZnGVs_boxaus_aug25.png", width = 20, height = 12, units = "cm")

 ggplot(ausZnGVs_definrange) +
  geom_histogram(aes(x=ZnPC95)) + 
  facet_wrap(~ State, scales = "free_y") +
   xlab(NULL) +
   ylab(expression("zinc BAPC95 (95% species protection, "*mu*"g/L)")) +
   theme_bw()

## New Zealand


znGV.summary <- nzZnGVs_definrange |>
  
  ungroup() |>
  filter(!is.na(ZnPC95)) |>
  group_by(Region) |>
  dplyr::summarise(
      Min = min(ZnPC95, na.rm = TRUE),
   Q5 = quantile(ZnPC95, probs = 0.05, na.rm = TRUE),
    Q25 = quantile(ZnPC95, probs = 0.25, na.rm = TRUE),
     Median = median(ZnPC95, na.rm = TRUE),
  Q75 = quantile(ZnPC95, probs = 0.75, na.rm = TRUE),
    Q95 = quantile(ZnPC95, probs = 0.95, na.rm = TRUE),
    Max = max(ZnPC95, na.rm = TRUE),
    N = n()
  ) |>
  arrange(Min)

znGV.summary |>
  flextable() |>
  autofit() |>
  theme_zebra()

    ggplot(nzZnGVs_definrange |> mutate(Region = ifelse(Region == "Manawatu-Whanganui",
                                                      "MW", Region))) +
  geom_boxplot(aes(x = reorder(Region, ZnPC95, FUN = median), y= ZnPC95)) + 
   xlab(NULL) +
   ylab(expression("zinc BAPC95 (95% species protection, "*mu*"g/L)")) +
   theme_bw()

# ggsave(file = "output/figuresnew/ZnGVs_boxnz_feb25.png", width = 20, height = 12, units = "cm")
ggsave(file = "data/BAGVs for Australia and New Zealand/output/ZnGVs_boxnz_aug25.png", width = 20, height = 12, units = "cm")

 
ggplot(nzZnGVs_definrange) +
  geom_histogram(aes(x=ZnPC95)) + 
  facet_wrap(~ Region, scales = "free_y") +
   xlab(NULL) +
   ylab(expression("zinc BAGV (95% species protection, "*mu*"g/L)")) +
   theme_bw()
```

### Density plots by State

```{r zndensity}

## Aussie
ZnQ5 <- quantile(ausZnGVs_definrange$ZnPC95, 0.05)

ZnQ5s <- ausZnGVs_definrange |> group_by(State) |> summarise(Line = quantile(ZnPC95, 0.05))

ggplot(data = ausZnGVs_definrange |> filter(!is.na(ZnPC95))) +
  geom_density(aes(x = ZnPC95))  +
  stat_summary(mapping = aes(x = ZnPC95, y = .25,
                   xintercept = ..x..),
              #fun = median,
              fun = function(z) { quantile(z,0.05) },
               geom = "vline",
               colour = "#011D77", orientation = "y") +
  geom_vline(aes(xintercept = ZnQ5), colour = "#333333", linetype = 2)+
  annotate("text", x=1.8*1.1, y=5.5, hjust = 0, label="National Q5", color = "#333333", size =3) +
  geom_text(data = ZnQ5s, aes(x=Line*1.1, y=4, hjust = 0), label="State Q5", color = "#011D77", size =3) +
  facet_wrap(~ State)+
 
  scale_x_log10() + 
  xlab(expression("zinc BAPC95 values ("*mu*"g/L)")) +
  ylab("Frequency") +
  theme_bw()

#ggsave(file = "output/figuresnew/StateQ5sZn_aus.png", width = 18, height = 12, units = "cm")
ggsave(file = "data/BAGVs for Australia and New Zealand/output/StateQ5sZn_aus_aug25.png", width = 18, height = 12, units = "cm")

## NZ
ZnQ5 <- quantile(nzZnGVs_definrange$ZnPC95, 0.05, na.rm=TRUE)

ZnQ5s <- nzZnGVs_definrange |> group_by(Region) |> summarise(Line = quantile(ZnPC95, 0.05, na.rm=TRUE))

ggplot(data = nzZnGVs_definrange |> filter(!is.na(ZnPC95))) +
  geom_density(aes(x = ZnPC95))  +
  stat_summary(mapping = aes(x = ZnPC95, y = .25,
                   xintercept = ..x..),
              #fun = median,
              fun = function(z) { quantile(z,0.05) },
               geom = "vline",
               colour = "#011D77", orientation = "y") +
  geom_vline(aes(xintercept = ZnQ5), colour = "#333333", linetype = 2)+
  annotate("text", x=0.8*1.1, y=4, hjust = 0, label="National Q5", color = "#333333", size =3) +
  geom_text(data = ZnQ5s, aes(x=Line*1.1, y=3, hjust = 0), label="Regional Q5", color = "#011D77", size =3) +
  facet_wrap(~ Region) +
 
  scale_x_log10() + 
    xlab(expression("zinc BAPC95 values ("*mu*"g/L)")) +
  ylab("Frequency") +
  theme_bw()

# ggsave(file = "output/figuresnew/Q5sZn_nz_region-feb.png", width = 18, height = 12, units = "cm")
ggsave(file = "data/BAGVs for Australia and New Zealand/output/Q5sZn_nz_region-aug25.png", width = 18, height = 12, units = "cm")
```

### CDFs

CDFs by State

```{r zncdfs}
###CDFs by State

plot_labs <- 
    ausZnGVs_definrange %>%
    group_by(State) %>%
    summarise(#ndata = tally(),
        ndata1 = n(),
        Q5 = unname(quantile(ZnPC95, probs=0.05, na.rm = TRUE))) |>
  mutate(State_labs = paste0(State, " n=", ndata1)
         )
        
ausZnGVs_definrange <- left_join(ausZnGVs_definrange, plot_labs, by = "State")


ggplot(ausZnGVs_definrange, aes(ZnPC95)) +
  stat_ecdf(geom = "step") + 
  facet_wrap(~ State_labs) +
  geom_vline(data = plot_labs, aes(xintercept = Q5),
             colour = "#FF931E", lty =2) +
  geom_text(data = plot_labs, aes(x = Q5*1.1, y = 0.95, label = format(round(Q5,1), nsmall = 1),
                                  hjust = 0), colour = "darkorange3") +
  theme_bw() +
  xlab(expression("zinc BAPC95 values ("*mu*"g/L)")) 

# ggsave(file = "output/figuresnew/Zn_StatesCDF_aus.png", width = 8, height = 5.6)
ggsave(file = "data/BAGVs for Australia and New Zealand/output/Zn_StatesCDF_aus_aug25.png", width = 8, height = 5.6)


## CDFs by region

plot_labs <- 
    nzZnGVs_definrange %>%
    group_by(Region) %>%
    summarise(#ndata = tally(),
        ndata1 = n(),
        Q5 = unname(quantile(ZnPC95, probs=0.05, na.rm = TRUE))) |>
  mutate(State_labs = paste0(Region, " n=", ndata1)
         )
        
nzZnGVs_definrange2 <- left_join(nzZnGVs_definrange, plot_labs, by = "Region")


ggplot(nzZnGVs_definrange2, aes(ZnPC95)) +
  stat_ecdf(geom = "step") +
  geom_vline(data = plot_labs, aes(xintercept = Q5),
             colour = "#FF931E", lty =2) +
  geom_text(data = plot_labs, aes(x = Q5*1.1, y = 0.95, label = format(round(Q5,1), nsmall = 1),
                                  hjust = 0), colour = "darkorange3") +
  theme_bw() +
  xlab(expression("zinc BAPC95 values ("*mu*"g/L)")) +
  facet_wrap(~ State_labs)

# ggsave(file = "output/figuresnew/Zn_StatesCDF_nz.png", width = 8, height = 5.6)
ggsave(file = "data/BAGVs for Australia and New Zealand/output/Zn_StatesCDF_nz_aug25.png", width = 8, height = 5.6)

```

## Summarise & weight by site

```{r site_summs_weights}
### Aussie
persite <- ausZnGVs_definrange |>
  group_by(State, Station, Stationme) |>
  summarise(Ndata = n())

totalsamples <- length(ausZnGVs_definrange$ZnPC80)
print("Total number of samples for Zn =")
totalsamples

print("Check that =")
totalsamples_chk <- sum(persite$Ndata)
totalsamples_chk

persite <- persite |>
  mutate(Weighting1 = 1/Ndata)

ausZnGVs_definrange <- left_join(ausZnGVs_definrange, persite |> dplyr::select(-Ndata))
library(modi)

weighted.quantile(ausZnGVs_definrange$ZnPC95, ausZnGVs_definrange$Weighting1, prob = 0.05, plot = TRUE)

quantile(ausZnGVs_definrange$ZnPC95,prob = 0.05)

T1DGV_table <- ausZnGVs_definrange |> group_by(State) |>
  summarise(N= n(),
            WQ5_PC99 = weighted.quantile(ZnPC99, Weighting1, prob = 0.05, plot = FALSE),
            WQ5_PC95 = weighted.quantile(ZnPC95, Weighting1, prob = 0.05, plot = FALSE),
            WQ5_PC90 = weighted.quantile(ZnPC90, Weighting1, prob = 0.05, plot = FALSE),
            WQ5_PC80 = weighted.quantile(ZnPC80, Weighting1, prob = 0.05, plot = FALSE))
  
T1DGV_table |>
  arrange(WQ5_PC95) |>
  flextable() |>
  theme_zebra() 
  
## NZ  
persite_nz <- nzZnGVs_definrange |>
  group_by(Region, Site_name, Site_code) |>
  summarise(Ndata = n())

totalsamples <- length(nzZnGVs_definrange$ZnPC80)
print("Total number of samples for Zn =")
totalsamples

print("Check that =")
totalsamples_chk <- sum(persite_nz$Ndata)
totalsamples_chk

persite_nz <- persite_nz |>
  mutate(Weighting1 = 1/Ndata)

nzZnGVs_definrange <- left_join(nzZnGVs_definrange, persite_nz |> dplyr::select(-Ndata))
library(modi)

weighted.quantile(nzZnGVs_definrange$ZnPC95, nzZnGVs_definrange$Weighting1, prob = 0.05, plot = TRUE)

quantile(nzZnGVs_definrange$ZnPC95,prob = 0.05)

T1DGV_table <- nzZnGVs_definrange |> group_by(Region) |>
  summarise(N= n(),
            WQ5_PC99 = weighted.quantile(ZnPC99, Weighting1, prob = 0.05, plot = FALSE),
            WQ5_PC95 = weighted.quantile(ZnPC95, Weighting1, prob = 0.05, plot = FALSE),
            WQ5_PC90 = weighted.quantile(ZnPC90, Weighting1, prob = 0.05, plot = FALSE),
            WQ5_PC80 = weighted.quantile(ZnPC80, Weighting1, prob = 0.05, plot = FALSE))
  
T1DGV_table |>
  arrange(WQ5_PC95) |>
  flextable() |>
  theme_zebra() 
```
